---
title: "12 Days of Christmas PA"
author: "Chau Nguyen"
date: "2/23/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(glue)
library(stringr)
library(purrr)
```

### A Test Dataset

Run the code below to load up a dataset called `xmas` that contains the crucial 
information about the gifts in the song. We will use this dataset to test out
our functions as we work on them.

```{r}
xmas <- read.csv("https://www.dropbox.com/s/e584pryn8evm1gz/xmas.csv?dl=1")
```

## Step One:  Plurals

Note that the gifts are listed in singular: for example, on day five the
narrator receives "five golden rings", but the entry in the dataset for the gift
on day five simply says "ring".  

Using the skeleton of the `pluralize_gift()` function, complete the code so that 
the function takes a gift and returns the appropriate plural. 

**Important:**  You should **absolutely not** "hard-code" anything into this 
function; this function should work in general, not just for the items in the 12 
Days of Christmas.  For example, the word "rings" should not appear anywhere in 
the function.  I should be able to give it any gift and get back the plural of 
that gift.

```{r}
# Function that takes a noun and makes it plural
# Arguments -- gift -- A string or vector of strings
# Return -- A string or vector of strings with the pluralized words
 
pluralize_gift <- function(gift){
gift <- case_when(str_detect(gift, 'y$') ~ 
                  str_replace(gift, 'y$', 'ies'),
                  str_detect(gift, 'oo') ~ 
                  str_replace(gift, 'oo','ee'),
                  TRUE ~ str_replace(gift, '$','s'))

return(gift)

}

gift <- xmas$Gift.Item

pluralize_gift(gift)

```

## Step Two: Creating sentences

Write a function called `make_phrase` that takes as input the necessary information, 
and returns a phrase.  For example, 



```{r}
make_phrase <- function(num, num_word, item, verb, adjective, location) {
  
  ## Step 1: Replace NAs with blank strings
  verb <- str_replace_na(verb, replacement ='')
  adjective <- str_replace_na(adjective, replacement = '')
  location <- str_replace_na(location, replacement = '')
  
  ## Step 2: If the day is larger than 1, the items need pluralized! 
  if(num != 1) {
    item <- pluralize_gift(item)
  }
  
  
  ## Step 3: If the word starts with a vowel, add "an" to the beginning
  if(num_word == "one") {
    num_word <- str_replace(num_word, "one", "a")
  }
  
  ## Step 4:" Glue all of the pieces together! 
  str_squish(glue::glue("{num_word} {adjective} {item} {verb} {location}"))
  
}
```

```{r}
xmas2 <- xmas %>%
  mutate(day.num = as.character(english::english(Day)
                                ), 
    Full.Phrase = pmap_chr(
      list(num = Day,
           num_word = day.num,
           item = Gift.Item,
           verb = Verb,
           adjective = Adjective,
           location = Location),
      make_phrase
      )
  )


```


### Challenge
```{r}
sing_line <- function(dataset, line, phrase_col) {
  num_word <- dataset[line, 2]
  intro <- glue::glue("On the {num_word} day of Christmas, my true love gave to me")
  phrases <- dataset %>% 
  pull({{phrase_col}})
  phrases <- phrases[line:1]
if(line != 1){
    phrases <- str_replace(phrases, "a", "and a") 
}
  for(i in phrases[line]){
    phrases[line:1]
  }
str_c(glue::glue("{phrases}"))
print(intro, phrases)
}
```


### A Small Test

```{r}
xmas2 %>% 
sing_line(line = 2, phrase_col = Full.Phrase)
```

### A Larger Test

```{r}
map_chr(1:12, ~ sing_line(xmas2, .x, Full.Phrase)) %>%
  str_c(collapse = "\n") %>% 
  cat()

```



