---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  warning = FALSE,
  message = FALSE
)
```

# clust431

<!-- badges: start -->
<!-- badges: end -->

The goal of clust431 is to use different clustering methods that can be used on datasets to cluster observations.

## Installation

You can install the released version of clust431 from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("clust431")
```

```{r example}
library(clust431)
library(tidyverse)
```

```{r}
iris2 <- iris %>%
    select(-Species)
```

### em_clust

```{r, warning = FALSE}
set.seed(175)
res <- em_clust(iris2, 3)

res$cluster

head(res$probabilities)

res$cluster_means

res$covariances

res$iterations
```

The output of the __em_clust()__ function is a vector of cluster assignments for each of the observations in the dataframe.

__How accurate were the cluster assignments?__

```{r}
res2 <- res$cluster
iris2 <- iris2 %>% 
    cbind(res2)
```

```{r}
ggplot(iris2, aes(x = Sepal.Length, 
                  y = Petal.Length,
                  color = as.factor(res2))) + 
    geom_point()
```

Looking only at Sepal Length and Petal Length in the iris dataset, we can see that the cluster assignments from the __em_clust()__ function make sense. The observations assigned to cluster 3 have similar Sepal and Petal lengths, and the observations in cluster 1 are all very similar in terms of these two variables.
