---
title: "Lab 1 part 2"
author: "Chau Nguyen"
date: "4/1/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(infer)
library(lubridate)
set.seed(1)


oscars <- read_csv(here::here("Oscars-demographics-DFE.csv"))

oscars <- oscars %>% 
  mutate(combined = paste(movie, award)) %>% 
  distinct(combined, .keep_all = TRUE)
```

__Most Big 5 Awards__
14 different movies won the most big 5 awards. They each won 3.
```{r}
most_awards <- oscars %>%
  group_by(movie) %>%
  distinct(award) %>%
  count(movie)

most_awards %>%
  filter(n == max(most_awards$n))

```


__Most Common First Name__

Helen is the most common first name of Best actress winners.
```{r}
oscars %>% 
  filter(award == "Best Actress") %>%
  distinct(person) %>%
  pull(person) %>%
  str_extract("[:alpha:]+") %>%
  table() %>%
  sort(decreasing = TRUE) %>%
  head(1)


```

__Most Oscar Winners__
New York state is the location of origin that has produced the most Oscar winners. 
```{r}
state_country <- oscars %>% 
  select(birthplace, person) %>%
  mutate(birthplace = if_else(birthplace == 'New York City', 
                              'Ny', birthplace)) %>%
  pull(birthplace) 

state_country <- str_extract(state_country, "[^ ]+$")

state_country <- as.data.frame(state_country) 

state_country  %>% 
  mutate(state_country = if_else(state_country == 'City', 
                              'Ny', state_country)) %>%
  count(state_country) %>% 
  slice_max(n = 1, n)
```


__Age and Gender__

```{r, error = TRUE}
oscars_dates <- oscars %>%
  filter(award != "Best Director") %>%
  mutate(date_of_show = gsub(" ", "", paste(year_of_award, "-02-01"))) %>% #found how to eliminate whitespace here: https://stackoverflow.com/questions/62000584/r-paste-two-strings-without-space 
  mutate(
    Age_At_Winning = case_when(
      str_detect(date_of_birth,c("-189|-188|-187|-186")) ~ as.numeric(as_date(date_of_show) - as_date(date_of_birth, format = "%d-%B-%Y"))/365.25, 
      TRUE ~ 100 + as.numeric(as_date(date_of_show) - as_date(date_of_birth, format = "%d-%B-%y"))/365.25)
         ) %>%
  mutate(Age_At_Winning = case_when(
    Age_At_Winning > 100 ~ Age_At_Winning - 100,
    TRUE ~ Age_At_Winning) #Adjustment made for winners born after the Unix Epoch and after who have their ages inflated by 100 years
  ) %>%
  mutate(
    Gender = case_when(
      str_detect(award, "Actor") ~ "Male",
      TRUE ~ "Female"
    )
  ) %>%
  mutate(
    year_of_award = year_of_award - min(year_of_award) #changing the 
  ) 


model <- lm(Age_At_Winning ~ year_of_award + Gender, data = oscars_dates)
summary(model)
  


```
```{r}
model
```

Our adjusted r-squared of .124 tells us the correlation is weak but an overall p-value of <.05 tells us our model is statistically significant. Looking at our predictor variables Gender had a p-value<.0001 which tells us their is strong confidence that gender is correlated with age of winner, with it estimating that men on average win big 5 awards ~9.14 years older than women do on average. The p-value of year_of_award was just slightly above the usual cutoff of .05 so we must reject it. If one were to test at a lower level of confidence they could conclude based on the model that with every decade that passes the award winners get on average ~0.55 years older. The intercept value of ~35.77 is what the model views as the expected average age of female winners during the first year the award show was held, 1929, and as it has a p-value<.05 we accept that this interpretation is statistically significant.

__Bootstrapping__

```{r}
set.seed(1) 
bootstrap_sample <- oscars %>%
  mutate(white_not_white = case_when(
    race_ethnicity == "White" ~ "White",
    TRUE ~ "Not White"
  )) %>%
  specify(response = white_not_white, success = "White") %>%
  generate(reps = 10000, type= "bootstrap") %>%
  group_by(white_not_white, replicate) %>%
  summarise(cnt = n()) %>%
  mutate(freq = cnt / 416, 3) %>%
  filter(white_not_white == "Not White")

```

```{r}

a = quantile(bootstrap_sample$freq, .025)
b = quantile(bootstrap_sample$freq, .975)

bootstrap_sample %>%
  ggplot(aes(x = freq)) +
  geom_histogram(aes(y = ..density..), bins = 30, fill = "grey", color = "black") + 
  geom_vline(xintercept = a, color = "blue") + 
  geom_vline(xintercept = b, color = "blue") + 
  ggtitle("Bootstrapped Plot") + xlab("Proportions")

```

```{r}
round(quantile(bootstrap_sample$freq, c(.025, .975)),3)

```

Through bootstrapping one can be 95% confident that the true proportion of non-white ‘Big 5’ award-winners is between 4.6% and 9.4%. We elected to include individuals who won multiple awards as multiple observations.