---
title: "Lab 7"
author: "Chau Nguyen"
date: "2/16/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

fish <- read_csv(here::here("Week 7","BlackfootFish.csv"))
```

## Summarizing Missing Data
Weight is the only variable that has missing values. Additionally, there are 1796 observations that have missing values.

```{r missing data, echo = FALSE}
find_na <- function(x){
  sum(is.na(x))
}

fish %>% 
  summarize(across(trip:species,  
                  ~ find_na(.x)))
```

# Task 2 – Visualizing Missing Data

Unfortunately, these missing values are not for only one year, trip, or section of river.

Create a thoughtful visualization exploring the frequency of missing values across the different years, sections, and trips.

```{r plot na, echo=FALSE}
fish_na <- fish %>% 
  group_by(year, section, trip) %>%
  summarize(missing = sum(find_na(weight)))

fish_na %>% 
  ggplot(mapping = aes(x = year, 
                       y = missing, 
                       fill = section)) +
  geom_col() +
  facet_wrap(~trip) +
  labs(y = "Missing Values", x = "Year") +
  theme(axis.text.x = element_text(size = 7), 
        axis.text.y =)
  
```

#Task 1 – Writing a Function
Let’s transform the repeated process above into a rescale_01() function.

The function should take a single vector as its input.
The function should return the rescaled vector.
Think about the “efficiency” of your function. Are you calling the same function multiple times?
```{r new rescale function}
rescale_01 <- function(x){
  rng <- range(x, na.rm = TRUE)
  (x - rng[1]) / (rng[2] - rng[1])
}
  
```

Task 3 - Performing a Simple Test
First, test your function on a simple vector. Verify that the maximum of your rescaled vector is 1 and the minimum is 0!

```{r Task 3, echo=FALSE}

x <- c(1:25, NA)

rescale_01(x)
```
Task 4 - Performing a More Difficult Test
Next, let’s test the function on the length column of the BlackfootFish dataset.

Make side-by-side plots of the original values of length and the rescaled values of length.
Hint: The gridExtra package has a grid.arrange() function that allows for you to arrange multiple ggplot() objects into rows and columns. Or, you can play around with the fig.show and out.width options of your code chunk (https://bookdown.org/yihui/rmarkdown-cookbook/figures-side.html).

```{r task 4, echo=FALSE, fig.show="hold", out.width="50%"}
fish_length <- fish %>% 
  pull(length)

rescaled_length <- rescale_01(fish_length)

fish$length_rescaled <- rescaled_length

fish %>% 
  ggplot(mapping = aes(x = year, 
                       y = length_rescaled))+
  geom_violin() +
  labs(x = "Year", y = "Rescaled Length of Fish")

fish %>% 
  ggplot(mapping = aes(x = year, 
                       y = length)) +
  geom_violin() +
  labs(x = "Year", y = "Length of Fish")

```
Task 5 – Incorportaing Variables
Suppose you would for your rescale() function to perform operations on a variable within a dataset. Ideally, your function would take a dataframe and a variable name as inputs and return a dataframe where the variable has been rescaled.

Create a rescale_column() that accepts two arguments, (1) a dataframe, and (2) the name(s) of the variable(s) to be rescaled. The body of the function should call the original rescale_01() function you wrote previously.

```{r Task 5, echo=FALSE}
rescale_column <- function(dataframe, vars){
  stopifnot(is.data.frame(dataframe))

dataframe %>% 
  mutate(
    across(.cols = {{vars}},
           .fns = rescale_01))
}
```

Task 6 – Another Function Test
Alright, now let’s put your rescale_column() function to work! Use your rescale_column() function to rescale both the length and weight columns.

I expect that you only call the rescale() function one time!
```{r Task 6, echo=FALSE}
fish %>% 
  rescale_column(length:weight)

```
```

